Fusion Moves for Correlation Clustering

### 2014-11-11 paper/egpaper_for_review.tex
update

% Include other packages here, before hyperref.
\usepackage{my_macros}
\usepackage{paralist}
\usepackage{amsthm}
% \usepackage{dirtytalk} fah: my miktex didn't like this; and we don't need it here. 
%\usepackage{framed}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{algorithm}
%\usepackage{algorithmic}
\usepackage{algpseudocode}
\usepackage{setspace}
%for tikz
\usepackage{tikz,pgfplots}
\usepackage{pgfplotstable}
\usepackage{filecontents}
%\usepackage{scrextend}
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

@@

%%%%%%%%% ABSTRACT
\begin{abstract}
  \color{red} \textbf{TODO:}
Correlation clustering, or multicut partitioning, is the NP-hard problem of partitioning an undirected graph or image with positive / attractive and negative / repulsive edge weights such that the sum of cut edge weights is minimized. We here present a move-making heuristic that uses cheaply generated proposal clusterings to iteratively and monotonously improve upon the best previously found clustering. A valid partitioning is maintained at all times, and solutions close to the global optimum are obtained faster than with any published technique.

%We address the problem of partitioning a  graph
%   into a previously unknown number of clusters.
%   Among all partitions, the one with the minimal 
%   sum of of cut edge weights is chosen. 
%   This problem is known as correlation clustering 
%   or multicut.
%   We propose a general framework to find
%   high quality approximate solutions for 
%   this NP-hard problem based on a move making algorithm:
%   We use candidate solutions from a proposal generator
%   to iteratively improve the best observed solution similar
%   to fusion moves.
%   The proposed solver outperforms any other solver
%   w.r.t. to any time performance.
%   The solutions found by this solver are close
%   to global optimal solutions w.r.t. energy
%   and problem specific measurements as VI for
%   image segmentation problems.

@@

Correlation clustering~\cite{Bansal-2002} also known as the multicut problem~\cite{chopra_1993_mp} 
is a basic primitive in computer vision~\cite{andres_2011_iccv,kroeger_2012_eccv,yarkony_2012_eccv,alush_2013_simbad} and data mining~\cite{Chierichetti-2014,Arasu-2009,Sadikov-2010,Chen-2012}. Its value is, firstly, that it accommodates both positive / attractive \emph{and} negative / repulsive edge weights. Importantly, this allows doing justice to evidence in the data that two nodes or pixels do \emph{not} wish to end up in the same cluster or segment. This is in contrast to the submodular potentials so popularized by the graph cut algorithm, which only allow two nodes to be attracted to each other, or at most be agnostic about membership in the same cluster. Secondly, the algorithm does not require a specification of the number of clusters. This is in contrast to methods such as normalized cut \cite{} that can only accommodate attractive interactions and hence need the number of clusters to be specified.

\paragraph{Contribution.} The evident usefulness of correlation clustering and its clean and compact formulation in terms of an optimization problem, together with its unfortunate NP-hardness, are an invitation to develop fast approximate solvers. The basic idea of the move making algorithm proposed here is to maintain, at all times, a best current partitioning; and to iteratively improve it (as we show, monotonously) by considering diverse and cheaply generated proposal partitionings. Any two nodes that are in one cluster in both the current and the proposal partitioning are contracted. Edges to contracted nodes are equally combined and reweighted, and the correlation clustering problem is then solved on this reduced graph. We offer a polyhedral characterization of this strategy, and evaluate it in conjunction with two versatile proposal generators.   We conduct experiments on a broad range of data sets ranging from 2D and 3D segmentation problems to clustering in signed social networks. The results suggest that this simple move making algorithm is the fastest technique known today, reaching close to globally optimal solutions in a tenth or a hundredth of the time required by the most efficient exact solvers.

\paragraph{Related Work.}
A natural approach is to solve the integer linear programming problem directly. To this end, efficient separation procedures have been found \cite{} that allow to iteratively augment the set of constraints until a valid partitioning is found. 
Alternatively, it is possible to relax the integrality constraints of the integer linear programming formulation. Such an outer relaxation can be iteratively tightened. However, intermediate solutions are fractional and so rounding is required to obtain a valid partitioning. The latter approach works best on planar \cite{yarkony_2012_eccv} or almost planar \cite{yarkony-andres-2013} graphs.

Another line of work maintains valid partitionings throughout, either by using a cluster representation in terms of labels \cite{bagon_2011_arxiv} or in terms of cut edges \cite{kroeger_2014_cvpr}. The former suffers \cite{beier} from the degeneracy of a label presentation \cite{kappes}, while the latter is highly efficient only for planar graphs. Working on contracted nodes has previously been explored in \cite{kroeger_2013_miccai}.

Finally, there are heuristic methods such as ...

---------------END OF SUGGESTION ----------------------------------

@@

\textbf{Outline:} {\bf If we are short of space, the following can be ommited:}
In sec.~\ref{sec:problem_formulation} we will give a 
detailed problem definition where we introduce 
the correlation clustering objective.
In sec.~\ref{sec:related_work} we will 
discuss existing methods for correlation 
clustering and briefly explain the concept of fusion moves.
In sec: ~\ref{sec:cc_fm} we describe our proposed
method and show the properties of the algorithm.
In sec.~\ref{sec:exp} we show an evaluation
of the method on existing parameter  and discuss the effects of parameters.
Future work will be discussed in sec. \ref{sec:future} and
we will conclude in ~\ref{sec:conclusion}.
%-------------------------------------------------------------------------

@@

\subsubsection{Randomized HierarchicalClustering (RHC)}

To generate proposals cut we can use bottom up hierarchical clustering.
In each step we contract the edge with the highest weight.
Parallel edges are merged into single edges by summing their weights.
We stop when a certain number of nodes is reached, or the
largest edge weight among the not yet contracted edges is smaller then a certain threshold.
To get versatile proposals we either add noise 
to the edge weighs or permute a certain number of weights.

@@

\subsubsection{Randomized Watersheds (RWS)}

The edge weighted watershed algorithm \cite{meyer_2013}
with random seeds can be used to get 
cheap proposals. Instead of $n$ seeds distributed uniformly
over all nodes we use the following.
Draw $n/k$ random edges only the negative edges, 
and give the two nodes of each random edge different seeds.
Doing so, a random subset of negative edges is forced
to be cut within each proposal.
For additional randomness, noise can be added to
the edge weighs.

@@

\subsection{Social Networks}
A application for large scale correlation clustering are social networks.
We consider two of those networks from the Stanford Large Network Dataset Collection\footnote{\url{http://snap.stanford.edu/data/index.html}}.
Both networks are given by weighted directed graphs with edge weights $-1$ and $1$. 
%
The first network is called \emph{Epinions}. 
This is who-trust-whom online social network of a a general consumer review site Epinions.com. 
Each directed edge $a\to b$ indicated that user $a$ trusts  or does not trust user $b$ is the edge-weight is positive or negative, respectively.
The network contains $131828$ nodes and $841372$ edges from which $85.3\%$ are positively weighted.
%
The first network is called \emph{Slashdot}. 
Slashdot is a technology-related news website know for its specific user community. 
In 2002 Slashdot introduced the Slashdot Zoo feature which allows users to tag each other as friends or foes. 
The network was obtained in November 2008 and contains $77350$ nodes and $516575$ edges from which $76.73\%$ are positively weighted.

We consider the problem to cluster this graphs such that positively weighted edges link inside and negatively weighted between clusters.
In other words friends and people who trust each other should be in the same segment and foes and non-trusting people in different clusters.
% 
To compensate the high impact of nodes with high degree we can normalize the edge weights such that each person has the same impact to the overall network, by enforcing.
\begin{align}
  \sum_{i\to j \in E_{\to}} |w_{i\to j}| &= 1&\forall i\in V, deg^{\textrm{out}}(i)\geq 1 
\end{align}
We define the following energy function
\begin{align}
 J(y) &= \sum_{i\to j \in E^+_{\to}} y_{ij}\cdot w_{i \to j} +  \sum_{i\to j \in E^-_{\to}} (y_{ij}-1)\cdot w_{i \to j} \nonumber\\
      &= \sum_{ij \in E} y_{ij}\cdot \underbrace{(w_{i \to j}+w_{j \to i})}_{w_{ij}} + \textrm{const}
\end{align}
which is zero if the given partitioning does not violated any relation and larger otherwise.

\subsection{(3D) Image Segmentation}
To segment images or volumes into a previously
unknown number of clusters, correlation clustering
can be used.
Starting from a super-pixel / super-voxel segmentation
which defines a region adjacency graphs (RAG),
correlation clustering finds the cut with the lowest energy.
The energy is based on a likelihood of merging adjacent super-voxels.
Each edge has a probability to keep adjacent segments separate ($p(y_{ij} =1)$)
or to merge them ($p(y_{ij} = 0)$).
The energy function is defined as following:
\begin{align}
 J(y)  &= \sum_{ij \in E} y_{ij}\cdot \underbrace{  log\left( \frac{p(y_{ij} =0)}{p(y_{ij} =1)}\right) + log \frac{1-\beta}{\beta}  }_{w_{ij}}
\end{align}
Where $\beta$ is used as a prior.

For 2D images, we took the models from \cite{andres_2011_iccv}
which are defined on the images of the Berkeley Segmentation
Database \cite{martin_2001}.
For 3D volume segmentation we took the models from
Andres \etal \cite{kroeger_2012_eccv}.
We use models with the cube sizes $150^3$, $300^3$, $450^3$, $550^3$ and $900^3$.
Both, the 2D and 3D models are available in a recent graphical model
benchmark of Kappes \etal~\cite{kappes_2013_benchmark_cvpr}.
We use the same names for the datasets as in the benchmark,
\emph{image-seg} for the models from  \cite{andres_2011_iccv} and \emph{knott-150 - knott-550} and \emph{3d-seg}
for the models from \cite{kroeger_2012_eccv}.
The size of instances is given in Tab.~\ref{tab:instance_sizes}.

\begin{table}[H]
   \tiny
   \centering
   \caption{Datasets used for evaluation}
   \label{tab:instance_sizes}
   \begin{tabular}{llllc}
      \toprule
         Dataset          &     \#Nodes     & \#Edges  & \#Instances  & Planar        \\ 
      \midrule 
         image-seg        &     156-3764        & ???        & 100          & $\bullet$    \\ 
         knott-150        &     572-972         & ???        & 2            & $\times$ \\
         knott-300        &     3846 - 5896     & ???        & 2            & $\times$ \\
         knott-450        &     15150-17070     & ???        & 2            & $\times$ \\
         knott-550        &     27286 - 31249   & ???        & 2            & $\times$ \\
         3d-seg           &     101220   & ???        & 1            & $\times$ \\
         slashdot         &     ???             & ???        & ???          & $\times$ \\     
      \bottomrule
   \end{tabular}
\end{table}

\begin{table}[H]
   \tiny
   \centering
   \caption{Eval}
   \label{tab:eval}
   \begin{tabular}{lllllll}
      \toprule
         Dataset          &  image-seg &  knott-3d-150 &knott-3d-300 &knott-3d-450 &knott-3d-550 &3d-seg\\
      \midrule 
         PIVOT-BOEM       &      \\ 
         HC               &\\
         HC-CGC           &    \\
         CGC              &   \\
         KL               & \\
         FUSION           &  \\
         MC-R             &  \\   
         MC-I             &  \\  
         FusionCC-HC-MC   &  \\  
         FusionCC-HC-CGC  &  \\  
         FusionCC-WS-MC   &  \\  
         FusionCC-WS-CGC  &  \\      
      \bottomrule
   \end{tabular}
\end{table}

>>>>

% Include other packages here, before hyperref.
\usepackage{my_macros}
\usepackage{paralist}
\usepackage{amsthm}
\usepackage{dirtytalk}
%\usepackage{framed}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{algorithm}
%\usepackage{algorithmic}
\usepackage{algpseudocode}
\usepackage{setspace}
%for tikz
\usepackage{tikz,pgfplots}
\usepackage{pgfplotstable}
\usepackage{filecontents}
%\usepackage{scrextend}
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

@@

%%%%%%%%% ABSTRACT
\begin{abstract}
  \color{red} \textbf{TODO:}
   We address the problem of partitioning a  graph
   into a previously unknown number of clusters.
   Among all partitions, the one with the minimal 
   sum of of cut edge weights is chosen. 
   This problem is known as correlation clustering 
   or multicut.
   We propose a general framework to find
   high quality approximate solutions for 
   this NP-hard problem based on a move making algorithm:
   We use candidate solutions from a proposal generator
   to iteratively improve the best observed solution similar
   to fusion moves.
   The proposed solver outperforms any other solver
   w.r.t. to any time performance.
   The solutions found by this solver are close
   to global optimal solutions w.r.t. energy
   and problem specific measurements as VI for
   image segmentation problems.

@@

@@

\textbf{Outline:}
In sec.~\ref{sec:problem_formulation} we will give a 
detailed problem definition where we introduce 
the correlation clustering objective.
In sec.~\ref{sec:related_work} we will 
discuss existing methods for correlation 
clustering and briefly explain the concept of fusion moves.
In sec: ~\ref{sec:cc_fm} we describe our proposed
method and show the properties of the algorithm.
In sec.~\ref{sec:exp} we show an evaluation
of the method on existing parameter  and discuss the effects of parameters.
Future work will be discussed in sec. \ref{sec:future} and
we will conclude in ~\ref{sec:conclusion}.
%-------------------------------------------------------------------------

@@

\noindent \textbf{Randomized HierarchicalClustering (RHC):}
%\subsubsection{Randomized HierarchicalClustering (RHC)}
%
To generate proposals cut we can use bottom up hierarchical clustering.
In each step we contract the edge with the highest weight.
Parallel edges are merged into single edges by summing their weights.
We stop when a certain number of nodes is reached, or the
largest edge weight among the not yet contracted edges is smaller then a certain threshold.
To get versatile proposals we either add noise 
to the edge weighs or permute a certain number of weights.

@@

\noindent \textbf{Randomized Watersheds (RWS):}
%\subsubsection{Randomized Watersheds (RWS)}
%
The edge weighted watershed algorithm \cite{meyer_2013}
with random seeds can be used to get 
cheap proposals. Instead of $n$ seeds distributed uniformly
over all nodes we use the following.
Draw $n/k$ random edges only the negative edges, 
and give the two nodes of each random edge different seeds.
Doing so, a random subset of negative edges is forced
to be cut within each proposal.
For additional randomness, noise can be added to
the edge weighs.

@@

\subsection{Social Networks and Network Clustering}
A application for large scale correlation clustering are social networks.
We consider two of those networks from the Stanford Large Network Dataset Collection\footnote{\url{http://snap.stanford.edu/data/index.html}}.
Both networks are given by weighted directed graphs with edge weights $-1$ and $1$. 
%
The first network is called \emph{Epinions}. 
This is who-trust-whom online social network of a a general consumer review site Epinions.com. 
Each directed edge $a\to b$ indicated that user $a$ trusts  or does not trust user $b$ is the edge-weight is positive or negative, respectively.
The network contains $131828$ nodes and $841372$ edges from which $85.3\%$ are positively weighted.
%
The second network is called \emph{Slashdot}. 
Slashdot is a technology-related news website know for its specific user community. 
In 2002 Slashdot introduced the Slashdot Zoo feature which allows users to tag each other as friends or foes. 
The network was obtained in November 2008 and contains $77350$ nodes and $516575$ edges from which $76.73\%$ are positively weighted.

We consider the problem to cluster this graphs such that positively weighted edges ($E^+_{\to}$) link inside and negatively weighted edges ($E^-_{\to}$) between clusters.
In other words friends and people who trust each other should be in the same segment and foes and non-trusting people in different clusters.
% 
To compensate the high impact of nodes with high degree we can normalize the edge weights such that each person has the same impact to the overall network, by enforcing.
\begin{align}
  \sum_{i\to j \in E_{\to}} |w_{i\to j}| &= 1&\forall i\in V, deg^{\textrm{out}}(i)\geq 1 
\end{align}
We define the following energy function
\begin{align}
 J(y) &= \sum_{i\to j \in E^+_{\to}} y_{ij}\cdot w_{i \to j} +  \sum_{i\to j \in E^-_{\to}} (y_{ij}-1)\cdot w_{i \to j} \nonumber\\
      &= \sum_{ij \in E} y_{ij}\cdot \underbrace{(w_{i \to j}+w_{j \to i})}_{w_{ij}} + \textrm{const}
\end{align}
which is zero if the given partitioning does not violated any relation and larger otherwise.

{\color{red}\textbf{TODO: Discussion}}

As another example for network clustering we use the modularity-clustering models from~\cite{kappes_2014_benchmark_arxiv} which are small but fully connected.

{\color{red}\textbf{TODO: Discussion}}

\subsection{2D and 3D Image Segmentation}
To segment images or volumes into a previously
unknown number of clusters, correlation clustering
has been used~\cite{andres_2011_iccv,kroeger_2012_eccv}.

Starting from a super-pixel/-voxel segmentation,
correlation clustering finds the clustering with the lowest energy.
The energy is based on a likelihood of merging adjacent super-voxels.
Each edge has a probability to keep adjacent segments separate ($p(y_{ij} =1)$)
or to merge them ($p(y_{ij} = 0)$).
The energy function is defined as following:
\begin{align}
 J(y)  &= \sum_{ij \in E} y_{ij}\cdot \underbrace{  log\left( \frac{p(y_{ij} =0)}{p(y_{ij} =1)}\right) + log \frac{1-\beta}{\beta}  }_{w_{ij}}
\end{align}
Where $\beta$ is used as a prior~\cite{andres_2011_iccv}.

We use the public available benchmark instances from~\cite{kappes_2013_benchmark_cvpr,kappes_2014_benchmark_arxiv}.
For 2D images, we took the segmentation problems on the Berkeley Segmentation Database~\cite{martin_2001} called \emph{image-seg}~\cite{andres_2011_iccv,kappes_2013_benchmark_cvpr}.
For 3D volume segmentation we took the models \emph{knott-3d-150},\emph{-300} and \emph{-450} from~\cite{kroeger_2012_eccv,kappes_2014_benchmark_arxiv} as well as the large
instance from the \emph{3d-seg} model~\cite{andres_2011_iccv,kappes_2013_benchmark_cvpr}. These instances have underling cube sizes of  $150^3$, $300^3$, $450^3$, and $900^3$, respectively.
We also request larger instances from the authors of~\cite{kroeger_2012_eccv} which kindly provide us the dataset~\emph{knott-3d-550} with cube size  $550^3$.
More information of the size of instances is given in Tab.~\ref{tab:instance_sizes}.

{\color{red}\textbf{TODO: Discussion}}

\begin{table}[H]
   \tiny
   \centering
   \caption{Datasets used for evaluation}
   \label{tab:instance_sizes}
   \begin{tabular}{llllc}
      \toprule
         Dataset          &     \#Nodes     & \#Edges  & \#Instances  & Planar        \\ 
      \midrule 
         image-seg        &     156-3764        & ???              & 100          & $\bullet$    \\ 
         knott-150        &     572-972         & ???              & 2            & $\times$ \\
         knott-300        &     3846 - 5896     & ???              & 2            & $\times$ \\
         knott-450        &     15150-17070     & ???              & 2            & $\times$ \\
         knott-550        &     27286 - 31249   & ???              & 2            & $\times$ \\
         3d-seg           &     101220          & ???              & 1            & $\times$ \\
         socialnets       &     77350 - 131828  & 516575 - 841372  & 2            & $\times$ \\  
      \bottomrule
   \end{tabular}
\end{table}

\begin{table}[H]
   \tiny
   \centering
   \caption{Evaluation of Variation of Information (VI) for datasets with available ground-truth.}
   \label{tab:eval}
   \begin{tabular}{lllllll}
      \toprule
         Dataset          &  image-seg   & knott-3d-150&knott-3d-300&knott-3d-450 &knott-3d-550 &3d-seg\\
      \midrule 
         PIVOT-BOEM       &   $ 4.9633$  &    $ 2.9936$ &       --       &\\ 
         HC               &   $ 2.5967$  &    $ 1.5477$ &    $ 2.3513$   &    $ 2.9155$ \\
         HC-CGC           &   $ 2.5164$  &    $ 0.9052$ &    $ 1.7636$   &    $ 2.2256$ \\
         CGC              &   $ 2.5247$  &    $ 0.9267$ &    $ 1.8822$   &    $ 2.3104$\\
         KL               &   $ 2.6432$  &    $ 2.0648$ &    $ 4.1318$   &    $ 4.9270$\\
         FUSION           &      --      &    $ 2.8787$ &    $ 4.0744$   &    $ 4.6616$\\
         MC-R             &   $ 2.5471$  &    $ 0.9178$ &    $ 1.6369$   &    $ 2.8710$\\   
         MC-I             &   $ 2.5367$  & $\bf 0.9063$ & $\bf 1.6352$   & $\bf 2.0037$ \\  
         FusionCC-HC-MC   &   $ 2.5319$  &    $ 0.9629$ &    $ 1.6516$   &    $ 2.0801$\\  
         FusionCC-HC-CGC  &$\bf 2.4961$  &    $ 0.9679$ &    $ 1.7673$   &    $ 2.3809$ \\  
         FusionCC-WS-MC   &   $ 2.5340$  &    $ 0.9629$ &    $ 1.6742$   &    $ 2.0739$\\  
         FusionCC-WS-CGC  &   $ 2.5192$  &    $ 1.0585$ &    $ 2.1344$   &    $ 2.7487$ \\      
      \bottomrule
   \end{tabular}
\end{table}
